---
title: "Penguins"
author: "Yu-En"
date: "13/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, cache = TRUE, message = FALSE)
```

## Import


```{r import}
library(tidymodels)
library(corrr)
library(moments)
data("penguins")
head(penguins)
penguins %>% glimpse()
penguins %>% skimr::skim()
penguins <- penguins %>%
  select(-island) %>%
  drop_na()
```

## Split Data

```{r split_data}
set.seed(300)
split <- initial_split(penguins, prop = 0.75)
penguins_train <- training(split)
penguins_test <- testing(split)
folds_5 <- vfold_cv(penguins_train, v = 5, repeats = 2)
```

## Show Random Forest Engine

```{r}
show_engines("rand_forest")
show_model_info("rand_forest")
```

## Tune Parameters

### 1. Use default parameters in `parsnip`

`rf_spec` is a random forest model specification created with `parsnip`. I do not specify values for any parameters, resulting in using the default values. As always, I then fit the model on the training data. The default parameters are printed.

```{r model_def}
rf_spec <- 
  rand_forest(mode = "regression") %>%
  set_engine("randomForest")

model_default <-
  rf_spec %>%
  fit(body_mass_g~., data = penguins_train)

model_default
model_default %>% 
  predict(penguins_test) %>% 
  bind_cols(penguins_test) %>% 
  metrics(body_mass_g, .pred)
```

### 2. Use `tune` to tune `parsnip` model

#### 2.1. Manually provide values

```{r model_tune1, collapse=TRUE}
# Update model specification
rf_spec <-
  rf_spec %>%
  update(mtry = tune(), trees = tune())

# Create workflow
rf_workflow <-
  workflow() %>%
  add_variables(outcomes = body_mass_g, predictors = everything()) %>%
  add_model(rf_spec)

# Put parameters in a grid
manual_grid <-
  expand.grid(mtry = c(1, 3, 5), trees = c(500, 1000, 2000))

# Tune
set.seed(300)
manual_tune <-
  rf_workflow %>%
  tune_grid(resamples = folds_5, grid = manual_grid)

# Show all results
collect_metrics(manual_tune)

# Show the best one
show_best(manual_tune, n = 1)

# Finalise
manual_final <-
  finalize_workflow(rf_workflow, select_best(manual_tune)) %>%
  fit(penguins_train)

# Predict on testing data
manual_final %>% 
  predict(penguins_test) %>% 
  bind_cols(penguins_test) %>% 
  metrics(body_mass_g, .pred)
```

#### 2.2. Specify grid size for automatic generation

```{r model_tune2, collapse=TRUE}
set.seed(300)
random_tune <-
  rf_workflow %>%
  tune_grid(
    resamples = folds_5, grid = 5
  )
collect_metrics(random_tune)

show_best(random_tune, n = 1)

random_final <-
  finalize_workflow(rf_workflow, select_best(random_tune)) %>%
  fit(penguins_train)

random_final %>% 
  predict(penguins_test) %>% 
  bind_cols(penguins_test) %>% 
  metrics(body_mass_g, .pred)
```

### 3. Create parameter values with `dials`

```{r model_tune3, collapse=TRUE}
mtry()

mtry() %>% range_get()

mtry() %>% range_set(c(1, ncol(penguins_train) - 1))

mtry(c(1, ncol(penguins_train) - 1))

cost_complexity()

trees()

set.seed(300)
trees() %>% value_seq(n = 4)

trees() %>% value_seq(n = 5)

trees() %>% value_seq(n = 10)

set.seed(300)
trees() %>% value_sample(n = 4)

trees() %>% value_sample(n = 5)

trees() %>% value_sample(n = 10)

set.seed(300)
dials_regular <- grid_regular(
  mtry(c(1, ncol(penguins_train) - 1)),
  trees(),
  levels = 3
)
dials_regular

set.seed(300)
dials_random <- grid_random(
  mtry(c(1, ncol(penguins_train) - 1)),
  trees(),
  size = 6
)
dials_random

dials_regular_tune <-
  rf_workflow %>%
  tune_grid(
    resamples = folds_5, grid = dials_regular
  )
collect_metrics(dials_regular_tune)
show_best(dials_regular_tune, n = 1)

dials_regular_final <-
  finalize_workflow(rf_workflow, select_best(dials_regular_tune)) %>%
  fit(penguins_train)

dials_regular_final %>% 
  predict(penguins_test) %>% 
  bind_cols(penguins_test) %>% 
  metrics(body_mass_g, .pred)

dials_random_tune <-
  rf_workflow %>%
  tune_grid(
    resamples = folds_5, grid = dials_random
  )
collect_metrics(dials_random_tune)
show_best(dials_random_tune, n = 1)

dials_random_final <-
  finalize_workflow(rf_workflow, select_best(dials_random_tune)) %>%
  fit(penguins_train)

dials_random_final %>% 
  predict(penguins_test) %>% 
  bind_cols(penguins_test) %>% 
  metrics(body_mass_g, .pred)
```

---

### Additional Information

* [Dials, Tune, and Parsnip: Tidymodels’ Way to Create and Tune Model Parameters](https://towardsdatascience.com/dials-tune-and-parsnip-tidymodels-way-to-create-and-tune-model-parameters-c97ba31d6173)